---
title: "Meta-analysis of *H. pylori* prevalence and antibiotic resistance in Central Asia"
author: "K. Dauyey"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(meta)
library(metafor)
library(readr)
library(dplyr)
```

## Inputs

Aggregate figures transcribed from published Central Asian studies. These files
**are** tracked in the repository under `data/meta/` (no patient data);
provenance and known data-quality items are in `data/meta/README.md`.

```{r load}
prevalence <- read_csv(here("data", "meta", "CentralAsia_Hpylori_Prevalence.csv"),
                       show_col_types = FALSE)
resistance <- read_csv(here("data", "meta", "CentralAsia_Hpylori_Resistance.csv"),
                       show_col_types = FALSE)
```

## Prevalence of *H. pylori*

Freeman–Tukey double-arcsine transform, random-effects model. Given the mix of
diagnostic methods (serology, UBT, PCR, culture, histology), heterogeneity is
expected to be high; pooled estimates are read as broad regional indicators.

```{r prevalence}
meta_prev <- metaprop(
  event   = Positive,
  n       = N,
  data    = prevalence,
  studlab = paste0(Author_Year, " [", Country, " - ", Diagnostic_Method, "]"),
  sm      = "PFT",
  method  = "Inverse",
  random  = TRUE
)
summary(meta_prev)
forest(meta_prev, xlab = "H. pylori prevalence (%)", col.diamond = "blue",
       print.I2 = TRUE)
```

### Sensitivity: excluding serology-based studies

```{r prevalence-sensitivity}
prev_no_sero <- prevalence %>%
  filter(!grepl("serolog|ELISA|IgG", Diagnostic_Method, ignore.case = TRUE))

if (nrow(prev_no_sero) >= 2) {
  meta_prev_ns <- metaprop(event = Positive, n = N, data = prev_no_sero,
                           studlab = prev_no_sero$Author_Year,
                           sm = "PFT", method = "Inverse", random = TRUE)
  summary(meta_prev_ns)
} else {
  message("Not enough non-serology studies for a sensitivity analysis.")
}
```

## Clarithromycin resistance

```{r clarithromycin}
meta_cla <- metaprop(
  event   = round(`Clarithromycin_Res_%` * N_Isolates / 100),
  n       = N_Isolates,
  data    = resistance,
  studlab = Author_Year,
  sm      = "PFT",
  random  = TRUE
)
summary(meta_cla)
forest(meta_cla, xlab = "Clarithromycin resistance (%)", col.diamond = "firebrick")
```

## Metronidazole resistance

```{r metronidazole}
res_mtz <- resistance %>% filter(!is.na(`Metronidazole_Res_%`))
if (nrow(res_mtz) >= 2) {
  meta_mtz <- metaprop(
    event = round(`Metronidazole_Res_%` * N_Isolates / 100),
    n = N_Isolates, data = res_mtz, studlab = res_mtz$Author_Year,
    sm = "PFT", random = TRUE)
  summary(meta_mtz)
  forest(meta_mtz, xlab = "Metronidazole resistance (%)", col.diamond = "orange")
}
```

## Amoxicillin resistance

```{r amoxicillin}
res_amx <- resistance %>% filter(!is.na(`Amoxicillin_Res_%`))
if (nrow(res_amx) >= 2) {
  meta_amx <- metaprop(
    event = round(`Amoxicillin_Res_%` * N_Isolates / 100),
    n = N_Isolates, data = res_amx, studlab = res_amx$Author_Year,
    sm = "PFT", random = TRUE)
  summary(meta_amx)
  forest(meta_amx, xlab = "Amoxicillin resistance (%)", col.diamond = "forestgreen")
}
```

## Session info

```{r session}
sessionInfo()
```
